Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-11-18 special-issue-ci/Response to reviewers.tex

\title{Volunteers in the Clouds: an Architecture for Low Cost and
  Potentially Massive Distributed Evolutionary Computation: Response to Reviewers\' comments}

\noindent
Dear Sirs,\\

We really appreciate the opportunity you have given us for improving
our work. Following the reviewers' requests and suggestions we have
deeply changed the contents, and thus, the quality of the paper.

We would like to notify the editors that, thanks to changes and
aportations made to this new version of the paper there are two new
coauthors.

@@

\begin{quote}
- According the last paragraph of the introduction, ``Then, a detailed description of the proposed system is
presented in Section III, followed by the experiments and obtained results in Section IV.", but Section IV
does not only contains experimental result but also the full description of the second version of the
architecture.
\end{quote}

You are absolutely right. We have corrected that paragraph to include
the actual contents of the paper.

\begin{quote}
Summarizing the review, the authors should provide a convincing example showing the advantages of
their proposal. As presented, the paper is a proof of concept of a system that seems difficult to be used
by the researchers needing high computational resources to run an evolutionary algorithm to solve a realworld
problem.
\end{quote}

We have added a new experiment (Section V.A), a more complex and
taxing problem that also uses a web-worker architecture, in order to
prove how this can be 
used in real-life problems. In this case, we have achieved many more
simultaneous users, up to 25 in some minutes and more than 100 in a
single hour. All in all, more than 400 volunteers (identified by
unique IPs) have participated in
the experiment.

In fact, our architecture {\sf NodIO} is
not an exclusively browser-based one. Its pool-based architeture allows
any other kind of client, including local ones, to be incorporated to
the experiment, with the lightweight overhead of doing asynchronous
HTTP requests, which being local are indeed negligible.

Our framework, in fact, does not prevent the user to carry out any other kind of
experiments. If there is a very efficient locally run algorithm, it only
needs to be able to do HTTP requests to work alongside a {\em
  volunteer cloud}. What our framework ensures is that all work done
by volunteers can be integrated with local, cloud or other kind of
resources. This explanation has been also added as a paragraph in the
Introduction to the paper.

@@

\begin{quote}
-  an experiment with a more difficult problem (please use the well-known benchmark problems
described in literature).
\end{quote}

A experiment using the well known Rastrigin function has been
performed and analyzed. It is now Subsection V.A.

\begin{quote}
-  A comparison in similar conditions of the two implementations of the architecture (the algorithms
used in the experiments are slightly different, but this could lead to dramatic changes in
performance, and consequently, invalidate the comparison performed in the paper). This would
enable a measurement of the actual improvement in performance obtained by the use of web
workers (not that generated due to a different algorithm)
\end{quote}

This has been included as the last subsection of the paper previous to
the conclusions. The new
experiment has constant population, instead of using random
population. We have not found a big influence on the result.

@@

\begin{quote}
Regarding the organization of the paper, some important details regarding the improved framework
design named is described in the experimentation section. I would reorganize the contents to describe the
improved in detail in section 3, and focus on the experiments, its
results and the discussion in section 4.
\end{quote}

In fact, reorganizing the paper increased the chance of introducing
new issues in this new version. We have preferred to keep it the way
it is, although we thank you for the suggestion. We expect that the
changes made to the text make it clear enough anyway.

@@

\begin{quote}
Another issue is that some details are described twice, for instance,
the advantages of the use of web 
workers. Fixing this issue can reduce the paper size and improve its
readability.
\end{quote}

We have eliminated duplications in this new version. Thanks for
noticing it.

\begin{quote}
Regarding the presentation:
-  Figures should be readable even in gray-scale, so do not use red and blue to denote different
elements, use a different shade or texture.
-  The axis labels and legends in most of the figures are unreadable, please increase its font size.
-  There are some typos and phrases that need rewriting.
\end{quote}

We have rewritten and revised extensively the paper, and tried to find
a way (and failed) to make it more readable for reviewers. In fact,
they use different shades, so it should mostly work in grayscale. For
the time being and 
during the revision phase, we prefer to keep it that way, but we will
change it to grayscale if it is required for the final version. 
In fact, the labels are not so important, so increasing the font size
would in fact distract the attention from the figure itself, which is
the interesting information we want to convey with the graph. Once
again, we can do it if required for the final version.

@@

\section{Response to Reviewer \#3}

\begin{quote}
1. Contribution: The paper presents the idea of a web browser based distributed evolutionary computing.
the system called NodIO consists of a REST server (coded in Node.js) and clients performing
evolutionary algorithm coded in JavaScript and embedded in a web page. Although the idea is not
entirely new, the authors show how Web Workers (introduced in HTML5) and the mechanism of
automatic restart of computations on the client's side can improve the performance of the system. The
state of the art is comprehensive and provides good introduction into the subject.
\end{quote}

We thank this reviewer for these helpful comments.

\begin{quote}
2. Relevance: In my opinion the system described in the paper is relevant to "Computational Intelligence
Software" special issue. Thanks to both development of Internet technology and continuous increase in
computing power of electronic devices, including mobile devices, the web browser based distributed
computing can be a cheap alternative for grid and high-performance computing systems that require
high capital expenditures and consume a lot of energy. Evolutionary computing system based on the
architecture presented inthe paper can be potentially used for solving many different time consuming
problems, however, the authors limited their experiments to a simple "Trap" function that can be easily
solved on a desktop machine, what may not be convincing for a reader looking for an effective
optimization tool.
\end{quote}

@@

It is true that that function can be easily solved with any desktop
machine, but we can still do that {\em and} use a volunteer
computation framework, which will add performance to the desktop
machine. Our intention was more to prove what kind of performance can
be expected and how this performance is distributed, although we have
also proved in the last section how the performance achieved by a
volunteer computer system can, in fact, be better than a desktop, even
a powerful one. This has also been clarified in the introduction and
conclusions.

At any rate, we have also performed a new experiment with the
Rastrigin function, which in fact cannot be solved in a few minutes
(or seconds, as in the case of Trap) by a desktop system.

\begin{quote}
3. Correctness: The system works correctly - it is not very complex and the authors provide its source
code (this is a great plus!). Technical description of the system is also clear, however it lacks some
details like parameters of the machine used to determine baselines for the experiments or the format of
messages sent between the server and the clients. The only thing I noticed is that the authors used the
term "multiple CPUs" in the description of Web Worker API advantages (page 8) instead of "multi-core
CPUs". As far as I know, contemporary web browsers do not support
multiple CPUs architecture.
\end{quote}

This has been clarified in the new version of the paper, third
paragraph from the beginning of the section that explains the web
worker-based architecture. In fact, most browsers can use multi-core
CPUs running web workers as independent processes in each one of them.
The base system has been now described in the first paragraph of the
section on the experiments, it is a desktop Intel i7-4770.

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex

In general, this experiment confirms what was found in the others in
several aspects: first, there is a regularity in the distribution of
time assigned by volunteers to it, second, there is a good and certain
amount of volunteers that will become part of the distributed
computer, in some cases up to one hundred; also, in some cases there
will be users that will spend several hours, two orders of magnitude
more generations in this case than in the previous one. There is a
clear influence of the time to solution: a short time means that every
experiment will only gather a few volunteers, but also that they will
stay only for a short time. A long time will make users stay for
longer, but if it takes too long, they will eventually lose interest.

>>>>

In general, this experiment confirms what was found in the others in
several aspects: first, there is a regularity in the distribution of the
time assigned by volunteers to it, second, there is a good and certain
amount of volunteers that will become part of the distributed
computer, in some cases up to one hundred; also, in some cases there
will be users that will spend several hours, two orders of magnitude
more generations in this case than in the previous one. There is a
clear influence of the time to solution: a short time means that every
experiment will only gather a few volunteers, but also that they will
stay only for a short time. A long time will make users stay for
longer, but if it takes too long, they will eventually lose interest.
